Fail Faster

### 2025-03-18 paper/paper.tex
progress

  % in strict functionallanguages like OCaml and Scal
To address this question, we propose a two-pronged approach for designing
generator libraries that preserves their expressiveness while eliminating
unnecessary overhead.
First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used for
optimizing functional DSLs~\cn.
We apply this technique to generators, completely eliminating the overhead of
many common generator abstractions. For example, in unstaged generator
libraries, each call to monadic bind usually spends time constructing and
immediately executing a closure at run time. \jwc{Todo: joe --- cite some evidence for this}
Meanwhile, a staged approach can resolve the bind at compile time, enabling the
compiler to inline and eliminate these closures.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison. \iflater\hg{More words here for
parallelism?}\fi Together, these insights provide a recipe for improving existing generator libraries.

@@

\jwc{Ensure we have a signposting table-of-contents-ish few sentences here, if it's not covered by the contribs.}

>>>>

  % in strict functionallanguages like OCaml and Scal
  To address this question, we provide two key insights for designing generator libraries, which eliminates unnecessary overhead while preserving expressiveness.
  First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used for
optimizing functional DSLs~\cn.
We apply this technique to generators, completely eliminating the overhead of
many common generator abstractions. For example, in unstaged generator
libraries, each call to monadic bind usually spends time constructing and
immediately executing a closure at run time. \jwc{Todo: joe --- cite some evidence for this}
Meanwhile, a staged approach can resolve the bind at compile time, enabling the
compiler to inline and eliminate these closures.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison. \iflater\hg{More words here for
parallelism?}\fi Together, these insights provide a recipe for improving existing generator libraries.

@@

\jwc{Ensure we have a signposting table-of-contents-ish few sentences here, if it's not covered by the contribs.}\todo{Todo Thia}
